---
title: "Sales Analysis"
author: "Rashaad Mohammed Mirza"
date: "Summer 2024"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
subtitle: "2020-2023"
---

# Report

```{r libraries, eval = T, warning = F, include = F}
# Import necessary libraries
library(DBI)
library(RSQLite)
library(RMySQL)
library(dplyr)
library(kableExtra)
library(ggplot2)

```

```{r connection, eval = T, warning = F, include = F}
# Connect to freemysqlhosting.net MySQL database from R
db_host <- "sql5.freemysqlhosting.net"
db_name <- "sql5724225"
db_user <- "sql5724225"
db_pwd <- "5YeYbiBfvH"
db_port <- 3306

# Connect to remote server database
mysql_conn <-  dbConnect(RMySQL::MySQL(),
                      host = db_host,
                      dbname = db_name,
                      user = db_user,
                      password = db_pwd,
                      port = db_port)

```

## AQI: Top 5 Products by Total Revenue (2020-2023)

```{r q1, warning = F, echo = F}
top_products <- "
SELECT product, SUM(total_amount) AS total_revenue
FROM sales_facts
GROUP BY product
ORDER BY total_revenue DESC
LIMIT 5
"

top_products_data <- dbGetQuery(mysql_conn, top_products)

```

The table and chart below show the top five products that generated the most revenue over the period from 2020 to 2023. This analysis helps in identifying the best-performing products and can guide strategic decisions such as inventory management, marketing focus, and sales strategies.

```{r r1, warning = F, echo = F}
# Table
knitr::kable(top_products_data, caption = "Top 5 Products by Total Revenue (2020-2023)")

# Bar Chart for Top Five Products with the Most Sales
ggplot(top_products_data, aes(x = reorder(product, -total_revenue), y = total_revenue)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Top 5 Products by Total Revenue",
       subtitle = "2020-2023",
       x = "Product",
       y = "Total Revenue") +
  coord_flip()

```

## AQII: Total Revenue and Units Sold per Product per Quarter (2020-2023)

```{r q2, warning = F, echo = F}
revenue_units <- "
SELECT 
    product, 
    year, 
    quarter, 
    SUM(total_amount) AS total_revenue, 
    SUM(total_units) AS total_units_sold
FROM 
    sales_facts
GROUP BY 
    product, 
    year, 
    quarter
ORDER BY 
    product, 
    year, 
    quarter
"

revenue_units_data <- dbGetQuery(mysql_conn, revenue_units)

```

The following analysis provides a breakdown of total revenue and units sold for each product on a quarterly basis. This information is useful for understanding seasonal trends, identifying products with consistent performance, and recognizing quarters with exceptional sales performance.

### Total Revenue and Units Sold per Product per Quarter

The table below shows the total revenue and units sold per product for each quarter from 2020 to 2023.

```{r r2t, warning = F, echo = F}
knitr::kable(revenue_units_data, caption = "Total Revenue and Units Sold per Product per Quarter")
```

### Total Revenue per Product per Quarter

The bar chart below visualizes the total revenue for each product, broken down by quarter. This helps in identifying which products performed best in specific quarters.

```{r r2b, warning = F, echo = F}
# Bar Chart for Total Revenue per Quarter
ggplot(revenue_units_data, aes(x = interaction(year, quarter), y = total_revenue, fill = product)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Total Revenue per Product per Quarter",
       x = "Year-Quarter",
       y = "Total Revenue",
       fill = "Product") +
  scale_x_discrete(labels = function(x) gsub("\\.", " Q", x)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Total Units Sold per Product per Quarter

The line chart below illustrates the total units sold for each product per quarter, highlighting trends in product demand over time.

```{r r2l, warning = F, echo = F}
# Line Chart for Total Units Sold per Quarter
ggplot(revenue_units_data, aes(x = interaction(year, quarter), y = total_units_sold, color = product, group = product)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Total Units Sold per Product per Quarter",
       x = "Year-Quarter",
       y = "Total Units Sold",
       color = "Product") +
  scale_x_discrete(labels = function(x) gsub("\\.", " Q", x)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## AQIII: Total Revenue per Product per Country (2020-2023)

```{r q3, warning = F, echo = F}
revenue_country <- "
SELECT 
    product, 
    country, 
    SUM(total_amount) AS total_revenue
FROM 
    sales_facts
GROUP BY 
    product, 
    country
ORDER BY 
    product, 
    country
"

revenue_country_data <- dbGetQuery(mysql_conn, revenue_country)

```

The line graph below illustrates the revenue trends for various products across different countries. Each line represents a product, allowing you to compare performance across regions.

```{r r3, warning = F, echo = F}
ggplot(revenue_country_data, aes(x = country, y = total_revenue, group = product, color = product)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Total Revenue per Product per Country",
       x = "Country",
       y = "Total Revenue",
       color = "Product") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_brewer(palette = "Set1")

```

## AQIV: Average Sales per Sales Rep per Quarter (2020-2023)

```{r q4, warning = F, echo = F}
avg_sales <- "
SELECT 
    rID, 
    sur, 
    first, 
    year, 
    quarter, 
    AVG(total_amount) AS avg_sales
FROM 
    rep_facts
GROUP BY 
    rID, 
    sur, 
    first, 
    year, 
    quarter
ORDER BY 
    year, 
    quarter, 
    avg_sales DESC
"

avg_sales_data <- dbGetQuery(mysql_conn, avg_sales)

```

The table below shows the average sales amount per sales rep for each quarter from 2020 to 2023. This information helps to easily identify key sales reps and their average sales performance across different quarters.

```{r r4, warning = F, echo = F}
avg_sales_data %>%
  kable("html", caption = "Average Sales per Sales Rep per Quarter") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F) %>%
  column_spec(1, bold = TRUE, color = "white", background = "#3c8dbc") %>%
  column_spec(6, bold = TRUE, color = "white", background = "#00a65a")

```

```{r disconnect, warning = F, include = F}
# Disconnect from the database when done
dbDisconnect(mysql_conn)

```

# References:
# https://www.youtube.com/watch?v=H6no78vdNUM
# https://northeastern.instructure.com/courses/180750/modules
# https://www.geeksforgeeks.org/fact-constellation-in-data-warehouse-modelling/
# https://chatgpt.com/
# https://www.databricks.com/glossary/snowflake-schema#:~:text=A%20snowflake%20schema%20is%20a,data%20marts%2C%20and%20relational%20databases.
# https://community.fabric.microsoft.com/t5/Desktop/Same-Dimensions-multiple-Fact-tables-different-measures-millions/td-p/3212805#:~:text=Given%20your%20requirements%2C%20there%20are%20a%20few%20potential%20approaches%20to%20consider.&text=Creating%20a%20star%20schema%20with,a%20bridge%20table%20where%20necessary.
# https://stackoverflow.com/questions/32139596/cannot-allocate-a-new-connection-16-connections-already-opened-rmysql
# https://www.lucidchart.com/pages/tour [ERD]
# https://www.phpmyadmin.co/index.php [Console]